R Course Syllabus

### 2023-04-27 Draft Syllabus.Rmd
Finalizing lesson 6

* [Codecademy Lesson](https://www.codecademy.com/courses/learn-r/lessons/intro-visualization-ggplot2-r/exercises/review)
* [W3 Schools](https://www.w3schools.com/r/r_if_else.asp)
* [Functionals](https://adv-r.hadley.nz/functionals.html#map)

@@

1. What is a function?
    + When to use functions.
2. Writing functions in R.
3. When to write a function.

@@

### **Helpful Links**
```{r lesson6}

```

>>>>

* [Codecademy Lesson Pt1](https://www.codecademy.com/courses/learn-r/lessons/r-data-cleaning/exercises/introduction)
* [Codecademy Lesson Pt2](https://www.codecademy.com/courses/learn-r/lessons/intro-visualization-ggplot2-r/exercises/review)
* [W3 Schools](https://www.w3schools.com/r/r_if_else.asp)
* [Functionals](https://adv-r.hadley.nz/functionals.html#map)

@@

1. What is a function?
    + When to use functions.
2. Writing functions in R.
3. Function nuances
    + Default values
    + Returning values

@@

```{r l6_2_1}
# A function that multiplies any number or vector by 10
mult10 <- function(x) {
  x <- x * 10
  return(x)
}

mult10(3)
mult10(20:22)
```

```{r l6_2_2}
# A function that counts the number of letters in a word
letter_count <- function(x) {
  # Stop if more than one word is provided
  if (length(x) > 1) stop("Error: You must provide a charcter vector of length 1")
  # Split the string
  x <- strsplit(x, ".", perl = T)
  # Count the number of splits
  return(length(x[[1]]))
}

letter_count("Reggie")
letter_count("supercalifragilisticexpialidocious")
try(letter_count(c("RG", "CG")))
```

```{r l6_2_3}
# A function to reverse the order of elements in a vector
vec_reverse <- function(x) {
  # Get the length of the vector
  vec_length <- length(x)
  # Get the class of the vector to create the appropriate container for the new vector
  mode <- class(x)

  # If x has length less than or equal to 1, no reversing is necessary
  if (vec_length <= 1) return(x)
  # Otherwise we reverse
  else {
    # Create an empty vector
    new_vec <- vector(mode = mode, length = vec_length)
    # Loop over the elements of the vector x
    for (i in seq(vec_length)) {
      # Element i in the new vector is element vec_length - (i-1) in the original vector.
      # For example if the original vector has length 5 then element 1 in the new vector is element
      # 5 - (1-1) = 5 in the original vector
      new_vec[i] <- x[vec_length - (i -1)]
    }
  }
  # Return the new vector
  return(new_vec)
}

vec_reverse(1:10)
vec_reverse(c(TRUE, TRUE, FALSE, TRUE, FALSE))
vec_reverse(as.numeric())
```

As you can see, after I put this code in a function I can call it any number of times. Be careful about assigning new things to the same name as was used to define a function. If you do this, your function will cease to work. Names in R can only be assigned to one object. For example, see what happens if I overwrite letter_count:

```{r l6_2_4}
# Assign the number of letters to letter_count
letters
letter_count <- length(letters)

# Try to call the function letter_count
try(letter_count("Hello"))
```

R tells me that the function does not exist!

#### **Function Nuances**

##### **Notes on arguments**

Each of the functions I've written thus far has required one argument, but you can write functions without arguments in R:

```{r l6_3_1}
no_args <- function() {
  "This is a function without arguments"
}
```

When you call such a function, it simply executes the code inside:

```{r l6_3_2}
no_args()
```

But most often, you will want to have arguments for your functions. This is because you will want your functions to be flexible--able to work under different circumstances and with varied input. You've already seen how to create a function that requires arguments:

```{r l6_3_3}
has_args <- function(x, y) {
  args <- match.call()
  num <- length(paste(args)) - 1
  paste("This function has", num, "argument(s)")
}

has_args(x = 1, y = 2)
```

Such a function fails if the user does not provide the correct number of arguments.

```{r l6_3_4}
try(has_args(1, 2, 3))
try(mult10())
```

To prevent this from happening, you can provide default arguments. If the user of the function does not provide an value to an argument, R will use the default you've provided. Let's update mult10 so that if the user does not provide a number or vector to be multiplied, it just returns `NA`.

```{r l6_3_5}
mult10 <- function(x = NA) {
  x <- x * 10
  return(x)
}
mult10()
```

You can provide as many defaults as you have arguments. Use these to help yourself or those who will be using your code down the line.

##### **Returning values**

If you want R to return a value to you after it finishes running your code, you can tell it that in several ways. Here are two:

1) End your function with the object you want returned:

```{r l6_3_6}
div2 <- function(x) {
  x <- x/2
  x
}
div2(30)
```

2) End your function with the a `return()` statement:

```{r l6_3_7}
square <- function(x) {
  x <- x^2
  return(x)
}
square(5)
```

Either is okay, but I recommend you get in the habit of using `return()`. It is clearer what you are doing and it will allow you to write code that can exit in multiple places if necessary.

* [Codecademy Lesson Pt1](https://www.codecademy.com/courses/learn-r/lessons/r-aggregates/exercises/introduction)
* [Codecademy Lesson Pt2](https://www.codecademy.com/courses/learn-r/lessons/r-multiple-tables/exercises/introduction)
* [W3 Schools](https://www.w3schools.com/r/r_functions.asp)
* [Functions](https://adv-r.hadley.nz/functions.html#functions)
